---
title: "MATH 342: HW 9 Problem 6"
author: "Samuel Eure"
date: "4/11/2018"
output: pdf_document
fontsize: 13pt
geometry: margin=0.5in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Question 6

##Part a
To double check my handwritten answer:
```{r}
d1 = c(1.328 ,1.342 ,1.075 ,1.228 ,0.939 ,1.004 ,1.178 ,1.286)
d2 = c(1.323,1.322,1.073,1.233,0.934,1.019,1.184,1.304)
d = cbind(d1,d2)
colnames(d) = c('op1', 'op2')
d = data.frame(d)
t.test(d$op1, d$op2)
```

Which results in the same p-value and test statistic I arrived at.  

##Part b
I can easily create 95% confidence intervals for the mean value of each of the operators measurements by the following:
$$
\bar{X_1} +/-  F^{-1}_{t_7}(\frac{1-0.05}2)\times\frac{s_1}{\sqrt8}
$$
$$
\bar{X_2} +/- F^{-1}_{t_7}(\frac{1-0.05}2)\times\frac{s_2}{\sqrt8}
$$

For operator 1:
```{r}
V = qt(0.975,7)*sd(d$op1)/sqrt(8)
mean(d$op1) - V
mean(d$op1) + V
```
I am 95% confident that the mean measurement taken by operator 1 is within [1.045584,1.299416]

For operator 2:
```{r}
V = qt(0.975,7)*sd(d$op2)/sqrt(8)
mean(d$op2) - V
mean(d$op2) + V
```
I am 95% confident that the mean measurement taken by operator 2 is within [1.049031,1.298969].  

##Part c
Through for my two-sample t-test, I assumed the variances were the same, which was ultimately shown to be a decient assumption since the Welch t-test resulted in the same p value and t statsitic.  I also assume that the data points are normally distributed around a mean and that the data points were collected independently of each other (i.e. iid). 
